A Switching Planner for Combined Task and Observation Planning

### 2011-03-09 subarchitectures/planner.sa/trunk/src/latex/IJCAI-Workshop/intro.tex
New version for IJCAI workshop

This paper describes a {\em switching} domain independent planning
approach we have developed to address this challenge. 
%%
Our Planner is {\em continual} in the usual sense that plans are
adapted and rebuilt online in reaction to changes to the model of the
underlying problem and/or domain---e.g., when goals are
modified, or when the topological map is altered by a door being
closed.
%%
It is integrated on a mobile robot platform that continuously
deliberates in a stochastic dynamic environment in order to achieve
goals set by the user, and acquire knowledge about its surroundings.
%%
Our planner takes problem and domain descriptions expressed in a novel
extension of PPDDL~\cite{younes:etal:2005}, called {\em
Decision-Theoretic DTPDDL}, for modelling stochastic decision
problems that feature partial observability.  In this paper we
restrict our attention to problem models that correspond to
deterministic-action multi-goal-oriented POMDPs 
% \footnote{We note that
% POMDPs with stochastic actions can be compiled into equivalent
% deterministic-action POMDPs, where all the original action uncertainty
% is expressed in the starting-state
% distribution~\cite{ng:Jordan:2000}.}
in which all actions have
non-zero cost---i.e., an optimal policy can be formatted as a finite
horizon contingent plan. Moreover, we target problems of a size and
complexity that is challenging to state-of-the-art sequential
satisficing planners, and which are too large to be solved directly by
decision-theoretic (DT) systems.

>>>>

This paper describes a {\em switching} domain independent planning
approach we have developed to address this challenge. 
%%
Our Planner is {\em continual} in the usual sense that plans are
adapted and rebuilt online in reaction to changes to the model of the
underlying problem and/or domain---e.g., when goals are
modified, or when the topological map is altered by a door being
closed.
%%
It is integrated on a mobile robot platform that continuously
deliberates in a stochastic dynamic environment in order to achieve
goals set by the user, and acquire knowledge about its surroundings.
%%
Our planner takes problem and domain descriptions expressed in a novel
extension of PPDDL~\cite{younes:etal:2005}, called {\em
Decision-Theoretic DTPDDL}, for modelling stochastic decision problems
that feature partial observability.  In this paper we restrict our
attention to problem models that correspond to deterministic-action
multi-goal-oriented POMDPs in which all actions have non-zero
cost\footnote{We note that POMDPs with stochastic actions can be
compiled into equivalent deterministic-action POMDPs, where all the
original action uncertainty is expressed in the starting-state
distribution~\cite{ng:Jordan:2000}.} -- i.e., an optimal policy can be
formatted as a finite horizon contingent plan. Moreover, we target
problems of a size and complexity that is challenging to
state-of-the-art sequential satisficing planners, and which are too
large to be solved directly by decision-theoretic (DT) systems.

### 2011-03-09 subarchitectures/planner.sa/trunk/src/latex/IJCAI-Workshop/switching.tex
New version for IJCAI workshop

We describe our approach that switches between sequential and
contingent sessions. As a continual planning approach, it proceeds by
interleaving planning and execution in a deterministic-action POMDP
described in DTPDDL. During a sequential session, planning is
performed by a ``classical'' system,\footnote{That is, a planner
designed to solve fully observable deterministic tasks.}  and
execution proceeds according to the {\em trace} computed by that
system. Taking the form of a classical plan, the trace specifies a
sequence of POMDP actions that achieve the agent's objectives in a
deterministic approximation, i.e., {\em determinisation}, of the
problem at hand. More precisely, the trace is an interleaved sequence
of POMDP actions and {\em assumptive} actions. The latter correspond
to assumptions the planner makes about the truth value of propositions
-- e.g. that a box of cornflakes is located in the kitchen at the
third plan step. They are called {\em applicability} assumptions if
the trace includes an action $\action$ that is not applicable with
probability $1$ at the belief-state
\bstate\ that the system is projected to be in when \action\ is
scheduled for execution, i.e., $\exists\state\in\states\;
\bstate(\state) > 0$ and $\poss(\action)\not\subseteq\state$. By
scheduling
\action,  the serial planner makes an assumption about the
observability of the precondition $\poss(\action)$.

Our approach always begins with a sequential session. Non-assumptive
actions from the trace are executed in sequence until the
applicability of the next scheduled action is uncertain. We denote
that action by \switchAction.  A contingent session then begins that
tailors sensory processing by the agent to determine whether the
assumptions made in the trace hold. 
%%
We add {\em disconfirm} and {\em confirm} actions to the model, so
the session is encouraged to judge the assumptions made in the
trace. On execution of one of those actions, control is returned to
the sequential session that continues at the current%%{\em underlying}
belief-state.
%%
%% For each assumption we add a {\em
%% disconfirm} action to the POMDP whose execution is rewarding
%% (resp. costly) if the assumption is false (resp. true). We also add
%% one {\em confirm} action that is rewarding (resp. costly) to execute
%% if $\poss(\switchAction)$ is true (resp. false). If execution of the
%% contingent plan applies a disconfirm action, then a new sequential
%% session begins at the {\em underlying} belief-state. If the confirm
%% action is executed, the sequential session is resumed, and
%% \switchAction\ is executed. Otherwise, control rests in the continual
%% session.
%%
Because contingent planning is only practical in relatively small
POMDPs, contingent sessions plan in an abstract decision process
determined by the current trace. This abstraction is constructed by
first omitting all propositions that do not feature in the trace, and
by then iteratively refining the model while the result is of a
practicable size.

Finally, whether proceeding in a sequential or contingent session, our
continual planner maintains a factored representation of successive
belief-states by performing belief revision according to
Eq~\ref{eq:revision}. Moreover, our internal representation of the
underlying belief-state corresponds more-or-less directly to an
$(\pp{:init})$ declaration. Our approach uses that distribution: (1)
as the source of candidate determinisations for sequential planning,
(2) to determine when to switch, and (3) as a mechanism to guide
construction of an abstract process for contingent sessions.

%% PLEASE STICK A PARAGRAPH HERE, (SENSIBLY) REQUESTED BY MORITZ, ABOUT
%% HOW WE DO BELIEF REVISION ACROSS SWITCHING.

%% \Omit{
%%   Therefore, although our current implementation does not
%%   explicitly support it, our approach generalises to POMDPs with
%%   stochastic action models.
%% }

\subsection{Sequential Planning in POMDPs}

A sequential session uses a classical planner to compute a trace. The
latter encapsulates assumptions about: (1) the true underlying state,
(2) how execution will progress, and (3) the possibility of the agent
eventually holding strong beliefs about the truth values of specific
state propositions. Here we describe the deterministic planning
problem, derived from the DTPDDL model, that admits plans which
correspond to traces.

In a deterministic-action POMDP all the uncertainty in state and
action is expressed in the $(\pp{:init})$ declaration. Our approach
uses the structure of that, as it occurs in the problem description,
to define a set of state-assumptions available to sequential planning.
%%
Writing \#\ if the value of a proposition is unspecified, for
$\pp{DORA}$ we have the following assumptions:

\small
\begin{tabular}{cccc}
\hline
Probability & (is-in DORA)  & (is-in box)  & (is-in cup) \\
\hline
.24 & kitchen & office & office \\
.06 & kitchen & kitchen & office \\
.56 & kitchen & office & kitchen \\
.14 & kitchen & kitchen & kitchen \\
.7 & kitchen & \# &  kitchen\\
.3 & kitchen & \# & office \\
.8 & kitchen & office & \# \\
.2 & kitchen & kitchen & \# \\
1.0 & kitchen & \# & \# \\
\hline
\end{tabular}
\normalsize

\noindent An assumption corresponds to a {\em
relaxed} visitation of the root term of $(\pp{:init})$. In this
relaxed case, a conjunctive term is visited iff its atomic subterms
are visited, and zero or more of its immediate probabilistic subterms
are visited. The starting state, $\state_0$, for sequential planning
is an {\em abstract} state that captures the unique assumption which
has probability $1$. In $\pp{DORA}$, that is:

\[
\begin{array}{l}
\state_0 \equiv \{(=(\pp{is-in}~\pp{DORA})~\pp{kitchen}),\\
\;\;(=(\pp{is-in}~\pp{box})~\#), (=(\pp{is-in}~\pp{cup})~\#)\}.
\end{array}
\]

In order to make assumptions available to the sequential planner, we
add one {\em assumptive} action $\assumptiveS{i}$ to the problem for
each element, $\prob_i (T_i)$, of each probabilistic term from
$(\pp{:init})$. The physics of these actions is as
follows: $\assumptiveS{i}$ can be executed if no $\assumptiveS{j}$,
$j \neq i$, has been executed from the same probabilistic term, and,
either $(\pp{probabilistic}~..\prob_i~(T_i)..)$ is in the root
conjunct, or it occurs in $T_k$ for some executed
$\assumptiveS{k}$. Also, an assumption about a proposition
$\prop$ can only be made once, and must be sequenced before any POMDP
actions that have $\prop$ in their precondition or effect.
%%
In other words, an assumptive action
must be applied to make a truth assignment to \prop, and then that
truth assignment can be modified by the DTPDDL domain operators.
%%
%%
Executing $\assumptiveS{i}$ in $\state$ effects a transition to a
successor state $\state^{T_i}$ with probability $\prob_i$, and
$\state^\bot$ with probability $1 - \prob_i$. Here, $\state^{T_i}$ is
the union of $\state$ with atomic terms from $T_i$ with the
proposition denoting the undefined fact is deleted. For example, if
the action adds $(=(\pp{is-in}~\pp{box})~\pp{office})$, then it must
delete $(=(\pp{is-in}~\pp{box})~\#)$. State $\state^\bot$ is an added
sink, where $\forall\action$ $\reward(\state^\bot, \action) = 0$ and
$\transProb(\state^\bot, \action, \state^\bot) = 1$.
%%
Lastly, we have that if $\action$ has a precondition that is not true
with probability $\theta$, a given threshold, in the underlying belief
$\bstate$, then immediately before it is executed blindly the system
switches to a contingent session (described below).

%% For sequential planning the operators from the POMDP model are
%% available with semantics that accommodate the above abstraction as
%% follows: In the deterministic model a proposition $\prop$ can be
%% thought to have a ternary interpretation at a state \state, as either
%% {\em true}, written $\prop\in\state$, {\em false}, written
%% $\prop\not\in\state$, or {\em unspecified}, with a slight abuse of
%% notation written $\prop\#\in\state$. For example, in $\pp{DORA}$
%% $(=(\pp{is-in}~\pp{box})\pp{office})\#\in\state_0$. For POMDP actions
%% $\action\in\actions$ in the deterministic model, if
%% $\prop \in \poss(\action)$, or if $\prop$ is the subject of a positive
%% or negative effect of $\action$, then $\action$ is not applicable
%% in \state\ if $\prop\#\in\state$.

%% Addressing now the switching semantics of action execution. I
%% %% In detail, we halt the sequential
%% session at \action\ if $\exists
%% \state$ s.t. $\bstate(\state)>0$, 
%% $\prop\in\poss(\action)$ , and $\prop\not\in\state$. According to the
%% semantics of action execution there must be a $\state'$ s.t.
%% $\bstate(\state')>0$ and $\prop\in\state'$. Otherwise, \action\ must
%% not have been scheduled at \bstate.

%% then $\reward(\state,\action)=10$, and
%% otherwise if $\prop\not\in\state'$ then
%% $\reward(\state',\action)=-10$.

We now describe the optimisation criteria given to the classical
planner. Where $\prob_i$ is the probability that the $i^{th}$
sequenced action $\action_i$ does not transition to $\state^\bot$, we
define the value of a trace
$\state_0, \action_0, \state_1, \action_1,.., \state_N$ from the
deterministic model to equal:

\begin{equation}\label{eq:tracevalue}
V(\state_0, \action_0, \state_1, \action_1,.., \state_N) =  \prod_{i=1..N-1} \prob_i \sum_{i=1..N-1} \reward(\state_i, \action_i)
\end{equation}

\noindent The optimal trace given a plan, i.e., the sequence of non-assumptive
actions from the trace, therefore has value:

\[
V^* = \max_{\prob_1, .., \prob_n} \prod_{i=1..N-1} \prob_i \sum_{i=1..N-1}
\reward(\state_i, \action_i),
\]

\noindent which is equal to the maximal 
additive contribution a trace of the plan can make to
Eq~\ref{eq:expectedvalue}.  In goal directed problems, where the only
non-zero reward is received at the first transition to a goal state,
the Eq~\ref{eq:tracevalue} criteria gives us the behaviour sought
by \system{FFR$_a$}~\cite{yoon:etal:2007}. From the perspective of
gradient-based reinforcement learning systems, such as William's
REINFORCE and GPOMDP variants for planning~\cite{olivier:doug:2009},
under reasonable assumptions the optimal trace identifies a maximal
gradient step from a uniformly random soft-max policy. Finally, it is
worth clarifying that although the above discussion has been in terms
of a trace of length $N$, there is nothing in our approach that
artificially limits the length of candidate sequential plans.

\subsection{Decision-Theoretic Planning in Abstractions}

In a contingent session our switching planner solves a sensing problem
in an abstract processes defined in terms of the assumptive actions in
the current trace, and the $(\pp{:init})$ declaration that models the
underlying belief-state.

The focus for contingent planning is on sensing, we therefore give the
process in this session an augmented reward model.  This reflects the
value of performing sensing actions in the context of the trace
proposed by the preceding sequential deliberation. First, all rewards
from the original POMDP are retained. Then, for each $\assumptiveS{i}$
action scheduled by the current trace, we have a {\em dual}
$\assumptiveDT{i}$ so that $\forall\state\in\states$:

\[
\reward(\state, \assumptiveDT{i}) = \bigg\{ \begin{array}{ll}
\$(T_i) & \pp{if}~\;\;T_i \not\subseteq \state \\
\hat\$(T_i) & \pp{otherwise} \\
\end{array}
\]

\noindent where $\$(T_i)$ (resp. $\hat\$(T_i)$) is a positive
(negative) numeric quantity which captures the utility the agent
receives for correctly (incorrectly) rejecting an assumption. The
added dual actions are propositionally trivial -- i.e.,
$\forall\state\in\states$ $\poss(\assumptiveDT{i})
\subseteq \state$ and $\assumptiveDT{i}(\state) \equiv \state$.  
%%
In order to reduce the number of possible rejections we omit
assumptions that are not {\em active} with respect to the action,
\switchAction, whose scheduled execution switched the system into
the contingent session. An assumption $\assumptiveS{i}$ is active if
\switchAction\ is not applicable in the sequential model unless
$\assumptiveS{i}$ is scheduled in the trace prefix
to \switchAction. For example, if $\pp{DORA}$'s trace is:

\[
\begin{array}{l}
\actions^{\circ}(.8;(=(\pp{is-in}~\pp{box})\pp{office}));\\
\actions^{\circ}(.3;(=(\pp{is-in}~\pp{cup})\pp{kitchen}));\\
(\pp{look-for-object}~\pp{DORA}~\pp{box}~\pp{office});\\
(\pp{look-for-object}~\pp{DORA}~\pp{cup}~\pp{kitchen});\\
(\pp{report-is-in}~\pp{box}~\pp{office}); \\
(\pp{report-is-in}~\pp{cup}~\pp{kitchen})
\end{array}
\]

\noindent Taking the switching action \switchAction\ to be
$(\pp{look-for-object}~\pp{DORA}~\pp{box}~\pp{office})$, we have that
$\actions^{\circ}(.3;(=(\pp{is-in}~\pp{cup})\pp{kitchen}))$ is not
active, and therefore exclude
$\actions^{\bullet}(.3;(=(\pp{is-in}~\pp{cup})\pp{kitchen}))$ from the
POMDP posed to the contingent session.

We also include another propositionally trivial action
$\actions.\poss(\switchAction)$ with the reward property:

@@

For the purposes of successive switching, execution of either a dual
action of the form $\assumptiveDT{i}$, or the
$\actions.\poss(\switchAction)$, returns control to a sequential
session. If a dual is executed, then the sequential planner must
replan with $\bstate_0$ equal to the underlying
belief-state. Otherwise, if $\actions.\poss(\switchAction)$ is
executed, then the current sequential plan is executed until further
sensing is scheduled, or to completion. Lastly, we note that in our
project the rewards, i.e., functions
$\$,\hat\$:2^\propositions\to\Re$, for the POMDP posed to the
contingent session are sourced from a motivational component of our
robotic architecture. In this paper, for $\assumptiveDT{i}$ actions we
set $\$(x)$ to be a positive constant, and have $\hat\$(x)= - \$(x)(1
- \prob) / \prob$ where $\prob$ is the probability that $x$ is true. For
$\actions.\poss(\switchAction)$ actions we have $\hat\$(x)= -
\$(x)\prob/(1-\prob)$.\footnote{These values were chosen to encourage
  sensing.}

%%  We make no assumptions here about how they are
%% derived.

The starting belief-state for the abstract process is given in terms
of a {\em relaxed} $(\pp{:init})$ declaration. We first construct a
tree which only includes terms from the {\em relaxed} visitation
characterised by the active {\em assumptive} actions in the current
sequential plan. Figure \ref{fig:abstraction} shows the belief-state
of a more detailed version of our previous example. The diamonds
represent probabilistic terms of the state description while circles
represent conjunctive or atomic terms.

@@

  \caption{Generating the abstract belief space.}
\label{fig:abstraction}
\end{figure}

% \small
% \begin{tabtt}
% (\=:init (= (is-in DORA) kitchen) \+ \\
%        (probabilistic \=.8 (= (is-in box) office)  \\
% 		      \>.2 (= (is-in box) kitchen))) \\
% \end{tabtt}
% \normalsize

% \noindent and therefore, the belief-state $\bstate_0$ is:

% \small
% \begin{tabular}{cccc}
% \hline
% Probability & (is-in DORA)  & (is-in box)  & (is-in cup) \\
% \hline
% .8 & kitchen & office & \# \\
% .2 & kitchen & kitchen & \# \\
% \hline
% \end{tabular}
% \normalsize

 % With
% regards to abstraction, the contingent and serial sessions apply the
% same applicability conditions ---i.e., that each effect and condition
% $\prop$ is specified, $\prop\in\state$--- to actions from the DTPDDL
% problem description.

In the first step, we remove all facts that are not part of an
assumption (Fig. \ref{fig:abstraction-b}). At this point, the session
would proceed in an abstraction of the environment that does not
contain $\pp{place1}$, the $\pp{cup}$ or a $\pp{kitchen}$. 
%%
In a second step, we iteratively refine the relaxed declaration by
adding terms from the original statement of $(\pp{:init})$ while the
number of abstract states in $\bstate_0$ that occur with non-zero
probability according to that refined declaration remains of a
practicable size. In detail, for each excluded atomic term, we compute
the {\em entropy} of the active assumptions, {\em conditional} on the
corresponding proposition.  Loosely speaking, the lower the {\em
  conditional entropy} of an atom, the more information about the
assumptions the contingent planner expect to derive from it. Thus,
propositions that in the first place were excluded are iteratively
added to the problem posed to the conditional session in increasing
order according to that measure (Fig. \ref{fig:abstraction-c}).

%% In detail, for each excluded atomic term, we compute the {\em
%% entropy} of the corresponding proposition, {\em conditional} on the
%% active assumptions. The {\em conditional entropy} measure
%% is higher the more uncertain that proposition is given truth
%% assignments to propositions that are the subject of active
%% assumptions. Propositions that in the first place were excluded are
%% iteratively added to the problem posed to the conditional session in
%% increasing order according to that measure.

%% In particular, we compute the {\em conditional entropy} of
%% each excluded proposition given the active assumptions. Propositions
%% are candidates for addition in increasing order according to that
%% measure.

%% Of course, it is expensive to measure size exactly,
%% and in practise we estimate this to be $2^{|\propositions^\#|}$, where
%% $\propositions^\#$ is the set of propositions in the abstraction that
%% were not assigned the {\em unknown} value \#. Terms are proposed for
%% addition because they are probabilistic, and because they were omitted
%% in the {\em relaxed} visitation characterised by {\em assumptive}
%% actions in the sequential plan. In each iteration, we add one
%% candidate term with the largest number of parent terms already in the
%% candidate tree.

%%% Local Variables: 
%%% mode: latex
%%% TeX-master: "moritz_2011"
%%% End:

>>>>

We now describe our {\em switching} planning system that operates
according to the continual planning paradigm. The system {\em
switches} in the sense that planning and plan execution proceed in
interleaved sessions in which the {\em base planner} is either {\em
sequential} or {\em decision-theoretic}.
%%
The first session is sequential, and begins when a DTPDDL description
of the current problem and domain are posted to the system.  During a
sequential session a serial plan is computed that corresponds to one
execution-{\em trace} in the underlying decision-process. That trace
is a reward-giving sequence of process actions and {\em assumptive}
actions. Each {\em assumptive} action corresponds to an assertion
about the truth of a fact that is unknown at plan time -- e.g. that a
box of cornflakes is located on the corner bench in the kitchen. The
trace specifies a plan and characterises a {\em deterministic
approximation} (see~\cite{yoon:etal:2008}) of the underlying process
in which that plan is valuable. Traces are computed by a
cost-optimising classical planner which trades off action costs, goal
rewards, and determinacy. Execution of a trace proceeds according to
the process actions in the order that they appear in the trace. If,
according to the underlying belief-state, the outcome of the next
action scheduled for execution is not predetermined above a threshold
(here 95\%), then the system switches to a DT session.

Because online DT planning is impractical for the size of problem we
are interested in, DT sessions plan in a small abstract problem
defined in terms of the trace from the proceeding sequential
session. This abstract state-space is characterised by a limited
number of propositions, chosen due to their {\em relevance} to the
action whose scheduled execution triggered the switch to the DT
session.
%%
To allow the DT planner to judge assumptions from the trace, we add
{\em disconfirm} and {\em confirm} actions to the problem for each of
them. Those yield a relatively small reward/penalty if the
corresponding judgement is true/false. If a judgement action is
scheduled for execution, then the DT session is terminated, and
a new sequential session begins.

%% Because online DT planning is impractical for the size of problem we
%% are interested in, when a DT session begins we use the trace from the
%% preceeding sequential session to restrict the state space. The state space is characterised by a limited number of propositions,
%% chosen due to their {\em relevance} to the action whose scheduled
%% execution triggered the switch to the DT session.
%% %The state
%% %space is constructed by first removing all propositions (and related
%% %actions) that are not true or assumed true of states in the trace,
%% %then adding them back as long as the number of states does not exceed
%% %a size threshold\footnote{Our experiments are performed using a number
%% %of different threshold parameters.}  Propositions to add back are
%% %chosen using as a heuristic the entropy of the trace assumptions
%% %conditional on the candidate proposition. 
%% To allow the DT planner to
%% determine the truth of the assumptions in the trace, we add {\em
%% disconfirm} and {\em confirm} actions to the problem for each of
%% them. These actions yield a relatively small reward/penalty if the
%% corresponding judgement is true/false. If a judgement action is
%% scheduled for execution the DT session is terminated, and control is
%% returned to a sequential session.

Whatever the session type, our continual planner maintains a factored
representation of successive belief-states.
% by performing belief revision. 
As an internal representation of the $(\pp{:init})$
declaration, we keep a tree-shaped Bayesian network which gets updated
whenever an action is performed, or an observation received. That
belief-state representation is used: (1) as the source of candidate
determinisations for sequential planning, (2) in determining when to
switch to a DT session, and (3) as a mechanism to guide construction
of an abstract process for DT sessions.

\subsection{Sequential Sessions}

As we only consider deterministic-action POMDPs, all state
uncertainty is expressed in the $(\pp{:init})$ declaration. This
declaration is used by our approach to define the starting state for
sequential sessions, and the set of state-assumptions available to
sequential planning.  Writing \#\ if the value of a proposition is
unspecified, taking the $(\pp{:init})$ example from the previous
section, we have the following assumptions:

\tiny
\begin{tabular}{cccc}
\hline
Probability & (is-in R2D2)  & (is-in box)  & (is-in cup) \\
\hline
%% .24 & kitchen & office & office \\
%% .06 & kitchen & kitchen & office \\
%% .56 & kitchen & office & kitchen \\
%% .14 & kitchen & kitchen & kitchen \\
.7 & kitchen & \# &  kitchen\\
.3 & kitchen & \# & office \\
.8 & kitchen & office & \# \\
.2 & kitchen & kitchen & \# \\
1.0 & kitchen & \# & \# \\
\hline
\end{tabular}
\normalsize

\noindent Each assumption corresponds to one distinct {\em
relaxed} visitation of the root term. Here, a conjunctive term is
visited iff its atomic subterms are visited, and zero or one of its
immediate probabilistic subterms are visited. For a sequential
session, the starting state corresponds to an assumption (usually
unique) with probability $1$.\footnote{Without a loss of generality,
we assume conjunctive terms list atoms that are true with
probability 1 according to a visitation, and that actions do not
have negative preconditions.}  In our example, that starting state is:

\vspace{-1ex}

\tiny
\[
\begin{array}{l}
\state_0 \equiv \{(=(\pp{is-in}~\pp{R2D2})~\pp{kitchen}),\\
\;\;(=(\pp{is-in}~\pp{box})~\#), (=(\pp{is-in}~\pp{cup})~\#)\}.
\end{array}
\]
\normalsize

\vspace{-1ex}

To represent state assumptions we augment the problem posed during
a sequential session with an \emph{assumptive action} $\assumptiveS{i}$ for
each element $\prob_i (T_i)$, of each probabilistic term from
$(\pp{:init})$. Here, $\assumptiveS{i}$ can be executed if no
$\assumptiveS{j}$, $j \neq i$, has been executed from the same
probabilistic term, and, either
$(\pp{probabilistic}~..\prob_i~(T_i)..)$ is in the root conjunct, or
it occurs in $T_k$ for some executed $\assumptiveS{k}$.
%%
We also add constraints that forbid scheduling of
assumptions about facts after actions with preconditions or effects
that mention those facts. For example, the robot cannot assume it is
plugged into a power source immediately after it unplugs itself.
%%
Executing $\assumptiveS{i}$ in a state $\state$ effects a transition
to a successor state $\state^{T_i}$ with probability $\prob_i$, and
$\state^\bot$ with probability $1 - \prob_i$. Here in addition to
respecting the above constraints, $\state^{T_i}$ is the union of
$\state$ with atomic terms from $T_i$. State $\state^\bot$ is an added
sink.

We now describe the optimisation criteria used during sequential
sessions. Let $\prob_i$ be the probability that the $i^{th}$ sequenced
action, $\action_i$, from a trace of state-action pairs
$\langle \state_0, \action_0,\state_1, \action_1,.., \state_N \rangle$
does not transition to $\state^\bot$, and
$\reward(\state_i, \action_i)$ be the instantaneous reward received
for executing action $\action_i$ in state $\state_i$. The optimal
trace has value:

%We now describe the optimisation criteria used during sequential
%sessions. Where $\prob_i$ is the probability that the $i^{th}$
%sequenced action, $\action_i$, from a trace of state-action pairs
%$\langle \state_0, \action_0,\state_1, \action_1,.., \state_N \rangle$
%does not transition to $\state^\bot$, we have that the optimal trace
%has value:

\tiny
\[
V^* = \max_N \max_{\state_0, \action_0,.., \state_N} \prod_{i=1..N-1} \prob_i \sum_{i=1..N-1}
\reward(\state_i, \action_i),
\]
\normalsize

%\noindent  Here,  $\reward(\state_i, \action_i)$ is the instantaneous
%reward ---i.e., can be negative for action costs, or positive for goal
%achievement--- received for executing action $\action_i$ in state
%$\state_i$. Finally, it is worth clarifying that in practise we do not
%artificially limit the length of traces that the sequential planner
%can consider. Moreover, for problems we consider there is always a
%finite optimal trace.

\subsection{DT Sessions}

During a sequential session, when an action is scheduled whose outcome
is uncertain according to the underlying belief-state, the planner
switches to a DT session. That plans for {\em small} abstract
processes defined according to the action that triggered the DT
session, the assumptive actions in the proceeding trace, and the
current belief-state. Targeted sensing is encouraged by augmenting the
reward model to reflect a heuristic value of knowing the truth about
assumptions. In detail, all rewards from the underlying problem are
retained. Additionally, for each relevant assumptive action
$\assumptiveS{i}$ in the current trace, we have a {\em disconfirm
action} $\assumptiveDT{i}$ so that for all states $\state$:

\vspace{-1ex}
\tiny
\[
\reward(\state, \assumptiveDT{i}) = \bigg\{ \begin{array}{ll}
\$(T_i) & \pp{if}~\;\;T_i \not\subseteq \state \\
\hat\$(T_i) & \pp{otherwise} \\
\end{array}
\]
\normalsize

\vspace{-1ex}

\noindent where $\$(T_i)$ (resp. $\hat\$(T_i)$) is a
small positive (negative) numeric quantity which captures
the utility the agent receives for correctly (incorrectly) rejecting
an assumption.
%%
In terms of action physics, a disconfirm action can only be executed
once, and otherwise is modelled as a self-transformation.
%%
We only consider {\em relevant} assumptions when constructing the
abstract model.  If \switchAction\ is the action that switched the
system to a DT session, then an assumption $\assumptiveS{i}$ is {\em
relevant} if it is necessary for the outcome of \switchAction\ to be
determined.  Continuing our simplified example, if the trace is:

\tiny
\[
\begin{array}{l}
\actions^{\circ}(.8;(=(\pp{is-in}~\pp{box})\pp{office}));\\
\actions^{\circ}(.3;(=(\pp{is-in}~\pp{cup})\pp{kitchen}));\\
(\pp{look}~\pp{box}~\pp{office});
(\pp{look}~\pp{cup}~\pp{kitchen});\\
(\pp{report}~\pp{box}~\pp{office}); 
(\pp{report}~\pp{cup}~\pp{kitchen})
\end{array}
\]
\normalsize

\noindent Taking the switching action \switchAction\ to be
$(\pp{look}~\pp{box}~\pp{office})$, we have that
$\actions^{\circ}(.3;(=(\pp{is-in}~\pp{cup})\pp{kitchen}))$ is not
relevant for this session, and therefore we exclude the corresponding
disconfirm action from the abstract decision process.

Given \switchAction, we also include another once-only self-transition
action $\actions.\poss(\switchAction)$, a \emph{confirmation action}
with the reward property:

@@

\noindent Here, \poss(\switchAction) is the set of positive literals
that comprise the precondition of \switchAction. Execution of
either a disconfirmation or the confirmation action returns control to
a sequential session, which then continues from the underlying
belief-state.

Turning to the detail of (dis-)confirmation rewards, in our integrated
system these are sourced from a motivational subsystem. In this paper,
for $\assumptiveDT{i}$ actions we set $\$(x)$ to be a small positive
constant, and have $\hat\$(x)= - \$(x)(1 - \prob) /
\prob$ where $\prob$ is the probability that $x$ is true. For
$\actions.\poss(\switchAction)$ actions we have $\hat\$(x)= -
\$(x)\prob/(1-\prob)$.

\Omit{
%%
Finally, 
If an
assumption was rejected, we prohibit that sequential session from
making it again.
}

%%  the size of the state space
%% is determined by two factors: the number of non-zero states in the
%% initial belief-state, and the number of actions. Determining
%% action reachability is intractable, as is determining their relevance
%% to goal achievement.

%% As we are aiming to keep the decision theoretic session fast, we try
%% to keep the state space of the abstract problem as small as
%% possible. In a deterministic-action POMDP, the state space size is
%% determined by two factors: the number of non-zero states in the
%% initial belief-state and the number of executable actions. As
%% determining the relevance of an action for a given goal in in general
%% a hard problem, we concentrate on keeping the number of uncertain
%% facts as low as possible without losing important information about
%% the sensing goal.

In order to guarantee fast DT sessions, planning occurs for a small
abstract problem. In our setting, where the underlying decision
process has a deterministic action model, a significant cause of
problem largeness is due to the number of states that occur with
non-zero probability in the underlying belief-state. We construct an
abstract belief-state that is chiefly characterised by facts related
to the current trace. In that abstraction, instances of operator or
sense declarations whose specification mentions an excluded fact are
not available.
%%
In detail, we first construct an $(\pp{:init})$ declaration which
contains the root term, and relevant assumptions from the trace---e.g., (Fig.\ref{fig:abstraction-b}) gives an example fragment, where
diamonds are probabilistic terms, and circles are atomic and/or
conjunctive.
%%
For each excluded atom, we compute the {\em entropy} of the {\em
relevant}
assumptions, {\em conditional} on that atom. Intuitively, lower
entropy indicates an atom gives better information about
assumptions. Facts are iteratively added to the belief-state in
increasing order according to that measure until the number of
abstract states in the initial belief reach predefined limit (Fig.\ref{fig:abstraction-c}).

%%  or no more atoms that reduce the entropy
%% can be added

\vspace{-.5cm}

@@

  \caption{Abstract belief-state, and refinement.}
\label{fig:abstraction}
\end{figure}

% When limiting the number of actions, we aim to include those that are
% possibly required to a) reach the goal or b) to directly observe an
% atom that is part of the abstract belief state. In general, checking
% this for an action is \textsc{PSPACE} complete for \textsc{STRIPS}
% planning.

% We approximate the set of relevant actions as follows: Let
% \relevantAtoms\ be the set of atoms that either occur in the goal
% condition \poss(\switchAction) or in a precondition \poss($a$) of an
% action $a$ that can directly sense any atom in the abstract belief
% state. We construct a \emph{relaxed planning graph} with
% \relevantAtoms\ as goal atoms from which we extract a relaxed
% plan.

%% In the first step, we remove all facts that are not part of an
%% assumption (Fig. \ref{fig:abstraction-b}). At this point, the session
%% would proceed in an abstraction of the environment that does not
%% contain $\pp{place1}$, the $\pp{cup}$ or a $\pp{kitchen}$. 
%% %%
%% In a second step, we iteratively refine the relaxed declaration by
%% adding terms from the original statement of $(\pp{:init})$ while the
%% number of abstract states in $\bstate_0$ that occur with non-zero
%% probability according to that refined declaration remains of a
%% practicable size. In detail,

%%% Local Variables: 
%%% mode: latex
%%% TeX-master: "aaai11"
%%% End:
